---
title: "regextable: Apply pattern-based text extraction and cleaning"
format: gfm
editor-options: 
  chunk-output-type: console
execute:
  eval: true   # do not run R code, just show it
  echo: true    # show the code in the output
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r}
#| include: false
devtools::load_all(".")
knitr::opts_chunk$set(
cache = FALSE,
collapse = FALSE,
warning = FALSE,
message = TRUE,
tidy = FALSE,
fig.align = "center",
comment = "#>",
fig.path = "man/figures/README-",
R.options = list(width = 200)
)
```

### Installation

```         
devtools::install_github("judgelord/regextable")
```

```{r, message=FALSE}
library(regextable)
```

## Data

This package operates on a data frame containing text to search and a corresponding regex lookup table containing patterns to match. Users must supply both the text data and the regex table, allowing extract() to work with any dataset or domain.

For example, you can provide a table of names, companies, or acronyms as the regex table, and any data frame of text as input. Example datasets `members` and `cr2007_03_01` from the legislators package are used here only for illustration. Users can supply any data frame and any regex table.

```{r}
data("members")
head(members)
data("cr2007_03_01")
head(cr2007_03_01)
```

## Text cleaning

Before searching text, `clean_text()` standardizes formatting to improve matching. It removes excess punctuation and spacing, lowercases text, and normalizes formatting.

```{r}
text <- "  HELLO---WORLD  "
cleaned_text <- clean_text(text)
print(cleaned_text)
```

## Extract regex-based matches from text

### Description

`extract()` performs regex-based matching on a text column using a lookup table. It returns a data frame with one row per detected pattern match and includes all original columns (or a subset if `return_cols` is specified).

### Required Parameters

-   **data**: A data frame or character vector containing the text to search.
-   **col_name**: Column name in the data frame containing text to search through.
-   **regex_table**: A regex lookup table with at least one pattern column.

### Optional Parameters

-   **pattern_col**: (default "pattern") Name of the regex pattern column in regex_table.
-   **return_cols**: (default NULL) Vector of columns to include in the output.
-   **id_col**: (default NULL) Column in data used for filtering rows before matching.
-   **id_filter**: (default NULL) Value(s) of IDs to restrict which rows are matched.
-   **date_col**: (default NULL) Column in data containing dates for filtering.
-   **date_start**: (default NULL) Start date for filtering rows.
-   **date_end**: (default NULL) End date for filtering rows.
-   **typo_table**: (default NULL) Table of typos to correct in data. (planned for future versions)
-   **remove_acronyms**: (default FALSE) If TRUE, removes all-uppercase patterns from regex_table.
-   **clean_text**: (default TRUE) If TRUE, applies basic text cleaning before matching.
-   **do_clean_text**: (default TRUE) Clean text before matching.
-   **verbose**: (default TRUE) If TRUE, displays progress messages.

### Returns

A data frame with one row per match, including: - All original columns from `data` (or `return_cols` if specified) - `pattern` — the first regex pattern matched in each row

### Basic Usage

The simplest use of `extract()` with only the required arguments. This finds all matches in the text column using the provided regex table.

```{r}
#Extract patterns using only required arguments
result <- extract(
  data = cr2007_03_01,
  col_name = "header",
  regex_table = members
)

head(result)
```

### Advanced Usage

Shows how to use optional arguments for more control, such as filtering by IDs, date ranges, and removing acronyms. Useful when you want to narrow the matches, not clean the text, or choose to display messages.

```{r}
# Advanced usage with optional filters
result_advanced <- extract(
  data = cr2007_03_01,
  col_name = "header",
  regex_table = members,
  id_filter = 80:130,         
  date_col = "date",               
  date_start = "2007-01-01",
  date_end = "2007-12-31",
  remove_acronyms = TRUE,
  return_cols = c("date", "header", "pattern")
)

head(result_advanced)
```

### Future Development

-   Add support for `typo_table` to correct known text errors before matching.
-   Improve strict matching rules for patterns that may need more inclusive or more restrictive word boundaries.\
-   Enable user-defined ID systems (e.g., corporations, campaigns) and control whether text is returned with matches.
