---
title: "regextable: Apply pattern-based text extraction and cleaning"
format: gfm
editor-options: 
  chunk-output-type: console
execute:
  eval: true
  echo: true
---

```{r}
#| include: false
devtools::load_all(".")
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  tidy = FALSE,
  fig.align = "center",
  comment = "#>",
  fig.path = "man/figures/README-",
  R.options = list(width = 120)
)
```
## Description 

`regextable` extracts regex-based pattern matches from a data frame or character vector using a lookup table of regular expressions. For each input row, the first matching pattern is returned along with the matched substring, an internal row identifier, and the original input columns (or a user-specified subset). Optional metadata from the pattern table can be included, and performance on large datasets is improved by stopping the search for a row once a match is found.

## Installation

```         
devtools::install_github("judgelord/regextable")
```

```{r, message=FALSE}
library(regextable)
```

## Data

This package operates on two inputs:

1.  A data frame containing the text to search.
2.  A regex lookup table containing patterns to match.

User-supplied patterns allow `extract()` to match text in any dataset or context.

The examples below use the `members` and `cr2007_03_01` datasets included in this package for illustration.

```{r}
data("members")
head(members)

data("cr2007_03_01")
head(cr2007_03_01)
```

## Text cleaning

Before matching, by default, `clean_text()` is applied to standardize text. It converts text to lowercase, removes excess punctuation, replaces line breaks and dashes with spaces, and collapses multiple spaces into a single space. Text cleaning is applied only during matching and does not modify the original input data. Users can disable this behavior by setting `do_clean_text = FALSE`.

```{r}
text <- "  HELLO---WORLD  "
cleaned_text <- clean_text(text)
print(cleaned_text)
```

## Extract regex-based matches from text

### Description

`extract()` performs regex-based matching on a text column using a pattern lookup table. For each input row, it returns at most one match (the first matched pattern), along with the corresponding pattern and optional metadata from the pattern table.

### Required Parameters

-   **`data`**: A data frame or character vector containing the text to search.
-   **`regex_table`**: A regex lookup table with at least one pattern column.

### Optional Parameters

-   **`col_name`**: (default `"text"`) Column name in the data frame containing text to search through.
-   **`pattern_col`**: (default `"pattern"`) Name of the regex pattern column in `regex_table`.
-   **`return_cols`**: (default `NULL`) Vector of columns from `data` to include in the output. If `NULL`, all columns from `data` are included.
-   **`regex_return_cols`**: (default `NULL`) Vector of additional columns from `regex_table` to include in the output.
-   **`date_col`**: (default `NULL`) Column in `data` containing dates for filtering.
-   **`date_start`**: (default `NULL`) Start date for filtering rows.
-   **`date_end`**: (default `NULL`) End date for filtering rows.
-   **`remove_acronyms`**: (default `FALSE`) If `TRUE`, removes all-uppercase patterns from `regex_table`.
-   **`do_clean_text`**: (default `TRUE`) If `TRUE`, cleans text before matching.
-   **`verbose`**: (default `TRUE`) If `TRUE`, displays progress messages.

### Returns

A data frame with one row per match, including:

- Original columns from data (or only `return_cols`, if specified)
- Additional columns from `regex_table` specified in `regex_return_cols`
- `pattern`, the first regex pattern matched in each row
- `row_id`, the row number of the text

### Basic Usage

The simplest use of `extract()` with only the required arguments and returned columns specified. This finds all matches in the text column using the provided regex table.

```{r}
#Extract patterns using only required arguments
result <- extract(
  data = cr2007_03_01,
  regex_table = members,
  return_cols = c("text"),
  regex_return_cols = c("icpsr") 
)

head(result)
```

### Advanced Usage

Shows how to use optional arguments for more control, such as filtering by date ranges and removing acronyms. This is useful when you want to narrow matches, disable text cleaning, control returned columns, or suppress messages.

```{r}
# Advanced usage with optional filters
result_advanced <- extract(
  data = cr2007_03_01,
  regex_table = members,
  date_col = "date",               
  date_start = "2007-01-01",
  date_end = "2007-12-31",
  remove_acronyms = TRUE,
  return_cols = c("text"),
  regex_return_cols = c("icpsr")
)

head(result_advanced)
```

### Future Development

-   Add support for `typo_table` to correct known text errors before matching.
-   Improve strict matching rules for patterns that may need more inclusive or more restrictive word boundaries.
-   Enable user-defined ID systems (e.g., corporations, campaigns) and control whether text is returned with matches.
