---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = FALSE,
  warning = FALSE,
  message = TRUE,
  tidy = FALSE,
  fig.align='center',
  comment = "#>",
  fig.path = "man/figures/README-",
  R.options = list(width = 200)
)
```

## regextable: Apply pattern-based text extraction and cleaning



### Installation
```
devtools::install_github("judgelord/regextable")
```

```{r, message=FALSE}
library(regextable)
```

## extract()

### Description
regextable provides tools for extracting regex-based pattern matches from text using a user-supplied lookup table. 
It is designed to work with any text dataset and any regex table, making it adaptable across domains 
(names, companies, acronyms, legislative text, etc.).

## Data
This package operates on a data frame containing text to search and a corresponding regex lookup table containing patterns to match. Users must supply both the text data and the regex table, allowing extract() to work with any dataset or domain.

For example, you can provide a table of names, companies, or acronyms as the regex table, and any data frame of text as input. Example datasets `members` and `cr2007_03_01` from the legislators package are used here only for illustration. Users can supply any data frame and any regex table.

```{r}
data("members")
head(members)
data("cr2007_03_01")
head(cr2007_03_01)
```

### Required Parameters
- **data**: A data frame or character vector containing the text to search.
- **col_name**: Column name in the data frame containing text to search through.
- **regex_table**: A regex lookup table with at least one pattern column.

### Optional Parameters
- **pattern_col**: (default "pattern") Name of the regex pattern column in regex_table.
- **return_cols**: (default NULL) Vector of columns to include in the output.
- **id_col**: (default NULL) Column in data used for filtering rows before matching.
- **id_filter**: (default NULL) Value(s) of IDs to restrict which rows are matched.
- **date_col**: (default NULL) Column in data containing dates for filtering.
- **date_start**: (default NULL) Start date for filtering rows.
- **date_end**: (default NULL) End date for filtering rows.
- **typo_table**: (default NULL) Table of typos to correct in data. (planned for future versions)
- **remove_acronyms**: (default FALSE) If TRUE, removes all-uppercase patterns from regex_table.
- **clean_text**: (default TRUE) If TRUE, applies basic text cleaning before matching.
- **do_clean_text**: (default TRUE) Clean text before matching.
- **verbose**: (default TRUE) If TRUE, displays progress messages.

### Returns
A data frame with one row per match, including:
- All original columns from `data` (or `return_cols` if specified)
- `pattern` — the first regex pattern matched in each row

### Basic Usage

```{r}
#Extract patterns using only required arguments
result <- extract(
  data = cr2007_03_01,
  col_name = "header",
  regex_table = members
)

head(result)
```

### Advanced Usage
```{r}
# Advanced usage with optional filters
result_advanced <- extract(
  data = cr2007_03_01,
  col_name = "header",
  regex_table = members,
  id_filter = 80:130,         
  date_col = "date",               
  date_start = "2007-01-01",
  date_end = "2007-12-31",
  remove_acronyms = TRUE,
  return_cols = c("date", "header", "pattern")
)

head(result_advanced)
```

### Future Development
- Add support for `typo_table` to correct known text errors before matching.
- Improve strict matching rules for patterns that may need more inclusive or more restrictive word boundaries.  
- Enable user-defined ID systems (e.g., corporations, campaigns) and control whether text is returned with matches. 