---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = FALSE,
  warning = FALSE,
  message = TRUE,
  tidy = FALSE,
  fig.align='center',
  comment = "#>",
  fig.path = "man/figures/README-",
  R.options = list(width = 200)
)
```

## regextable: Apply pattern-based text extraction and cleaning



### Installation
```
devtools::install_github("judgelord/regextable")
```

```{r, message=FALSE}
library(regextable)
```

## extract()

### Description
`extract()` is the core function of the regextable package. It uses a user-supplied regex lookup table to extract pattern matches from a data frame or character vector. The function returns a data frame containing the original columns plus a `pattern` column for each detected pattern or only user specified columns.

This package is generic: it works with any text data and any regex table. For example, you can extract names, companies, acronyms, or other entities from text. Users can supply their own data frames and pattern tables to adapt `extract()` to their needs.

## Data
This package operates on a data frame containing text to search and a corresponding regex lookup table containing patterns to match. Users must supply both the text data and the regex table, allowing extract() to work with any dataset or domain.

For example, you can provide a table of names, companies, or acronyms as the regex table, and any data frame of text as input. Example datasets `members` and `cr2007_03_01` from the legislators package are used here only for illustration. Users can supply any data frame and any regex table.

```{r}
data("members")
head(members)
data("cr2007_03_01")
head(cr2007_03_01)
```

### Required Parameters
- **data**: A data frame or character vector containing the text to search.
- **col_name**: Column name in the data frame containing text to search through.
- **regex_table**: A regex lookup table with at least one pattern column.

### Optional Parameters
- **pattern_col**: (default "pattern") Name of the regex pattern column in regex_table.
- **return_cols**: (default NULL) Vector of columns to include in the output.
- **id_col**: (default NULL) Column in data used for filtering rows before matching.
- **id_filter**: (default NULL) Value(s) of IDs to restrict which rows are matched.
- **date_col**: (default NULL) Column in data containing dates for filtering.
- **date_start**: (default NULL) Start date for filtering rows.
- **date_end**: (default NULL) End date for filtering rows.
- **typo_table**: (default NULL) Table of typos to correct in data. (planned for future versions)
- **remove_acronyms**: (default FALSE) If TRUE, removes all-uppercase patterns from regex_table.
- **clean_text**: (default TRUE) If TRUE, applies basic text cleaning before matching.
- **do_clean_text**: (default TRUE) Clean text before matching.
- **verbose**: (default TRUE) If TRUE, displays progress messages.

### Returns
A data frame with one row per match, including:
- All original columns from `data` (or `return_cols` if specified)
- `pattern` — the first regex pattern matched in each row

### Basic Usage

```{r}
#Extract patterns using only required arguments
result <- extract(
  data = cr2007_03_01,
  col_name = "header",
  regex_table = members
)

head(result)
```

### Advanced Usage
```{r}
# Advanced usage with optional filters
result_advanced <- extract(
  data = cr2007_03_01,
  col_name = "header",
  regex_table = members,
  id_filter = 80:130,         
  date_col = "date",               
  date_start = "2007-01-01",
  date_end = "2007-12-31",
  remove_acronyms = TRUE,
  return_cols = c("date", "header", "pattern")
)

head(result_advanced)
```


> in the future, `extractMemberName()` may default to returning a list of dataframes the same length as the supplied data

For example, we can use `extractMemberName()` to detect the names of members of Congress in the text of the Congressional Record. Let's start with text from the Congressional Record from 3/1/2007, scraped and parsed using methods described [here](https://judgelord.github.io/congressionalrecord/). 

```{r}
data("cr2007_03_01")

head(cr2007_03_01)

head(cr2007_03_01$url)
```

This is an extremely simple example because the text strings containing the names of the members of Congress in the `speaker` column are short, consistently formatted, and do not contain much other text. However, `extractMemberName()` can also search much longer and messier texts, including text where names are not consistently formatted or where they contain common typos introduced by humans or common OCR errors. Indeed, these functions were developed to identify members of Congress in ugly text data like [this](https://judgelord.github.io/corr/corr_pres.html#22). 

To better match member names, this function currently requires either:

- a column "congress" (this can be created from a date) or 
- a vector of congresses to limit the search to supplied to the `congresses` argument

For illustration, I show both options in the example below.

Member names augmented from voteview come with the `legislators` package, but users can also supply a customized version of these data to the `members` argument.

### `extractMemberName()`

```{r, message=TRUE}
# extract legislator names and match to voteview ICPSR numbers
cr <- extractMemberName(data = cr2007_03_01, 
                        col_name = "speaker", # The text strings to search
                        congress = 110,
                        cl = 4 #Use 4 cores (on Mac)
)

cr
```

In this example, all observations are in the 110th Congress, so we only search for members who served in the 110th.

Because `extractMemberName()` links each detected name to ICPSR IDs from voteview.com, we already have some information, like state and district for each legislator detected in the text (scroll to the right). 



### Future Development
- Add support for `typo_table` to correct known text errors before matching.
- Improve strict matching rules for patterns that may need more inclusive or more restrictive word boundaries.  
- Enable user-defined ID systems (e.g., corporations, campaigns) and control whether text is returned with matches. 

### Dynamic data 

- [ ] make some kind of version-controlled spreadsheet where users can submit additional nicknames, maiden names, etc. to augment the voteview data. This is currently done in data/make_members/nameCongress.R, but it would not be easy for users to edit and then re-make the regex table.
- [ ] make some sort of version-controlled spreadsheet where users can submit common typos they find. Currently, typos.R generates typos.rda, which is loaded with the package.
 
### Vignettes

- [ ] Add a vignette using messy OCRed legislator letters to FERC from replication data for ["Legislator Advocacy on Behalf of Constituents and Corporate Donors"](https://judgelord.github.io/research/ferc/)
- [ ] Add a vignette using [public comment data](https://github.com/judgelord/rulemaking) (sparse legislator names)

### Functions 

- [X] `extractMemberName()` needs options to supply custom typos and additions to the main regex table
- [X] correcting OCR errors and typos should be optional in `extractMemberName()`
- [ ] integrate with the [`congress`](https://github.com/ippsr/congress) and/or [`congressData`](https://github.com/IPPSR/congress) packages. For example, we may want a function (`augmentCongress` or `augment_legislators`?) to join in identifiers for other datasets on ICPSR numbers. Perhaps this is best left to users using the `congress` package. 
- [ ] Additionally, `committees.R` provides a crosswalk for Stewart ICPSR numbers
- [ ] Support dates for congress

### Documentation 

- [ ] Document helper functions for `extractMemberName()`
- [X] Document additional functions that help prep text for best matching
- [ ] Document example data (including FERC and public comment data for new vignettes)
