---
title: "Agencies Regex Table"
author: "Shirlyn Dong"
date: "`r Sys.Date()`"
format: 
  html:
    toc: true
vignette: >
  %\VignetteIndexEntry{Agencies}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = FALSE,
  warning = FALSE,
  message = FALSE,
  tidy = FALSE,
  fig.align = 'center',
  R.options = list(width = 120)
)

kable <- function(x, ...){
  if (knitr::is_latex_output()) {
    knitr::kable(x, booktabs = TRUE, format = 'latex', ...) |>
      kableExtra::kable_styling(latex_options = c("striped", "scale_down", "HOLD_position"))
  } else {
    knitr::kable(x, escape = TRUE, ...) |>
      kableExtra::kable_styling()
  }
}
```

### Introduction

This vignette demonstrates how to extract sub-agency names from texts using a regex lookup table that includes agencies, sub-agency names, sections, and departments. The `regextable::extract()` function provides a consistent workflow for identifying sub-agency names across different texts.

Install and load the package:
```{r, message=FALSE}
library(regextable)
library(kableExtra)
library(rvest)
library(googlesheets4)
library(purrr)
library(tibble)
library(stringr)
```

### Agencies Data File

The `agencies_by_section` table contains a sample of agencies, including their sub-agencies, acronyms, and reference sources used for matching and standardization. These serve as the regex patterns used for matching and standardization.

```{r}
load("/Users/shirl/Downloads/agencies_by_section.rda")
kable(head(agencies_by_section))
```

### Body Data Files

The `body_parsed` dataset contains text sections along with associated organizational metadata. Each row includes the full text, the section and department it belongs to, the agency and sub-agency names, sub-agency acronyms, and any detected mentions of agencies or acronyms.

```{r}
load("/Users/shirl/Downloads/body_parsed.rda")
kable(head(body_parsed))
```

### Extracting Sub-Agency Names from Body Texts

This example demonstrates how to search the `text` column of the `body_parsed` dataset using patterns from the `subagency_name` column of the `agencies_by_section` regex table.

By `setting use_ner = TRUE`, the `extract()` function performs Named Entity Recognition (NER) to validate that the regex matches actually correspond to organizational entities in the `text`. The final output returns the matched sub-1 names, retains the text and `department` columns from the original data, and attaches the corresponding `agency` metadata from the lookup table.

```{r}
agencies_df <- extract(
  data = body_parsed,
  col_name = "text",
  regex_table = agencies_by_section,
  pattern_col = "subagency_name",
  data_return_cols = c("text", "department"),
  regex_return_cols = c("agency"),
  use_ner = TRUE
)
kable(head(agencies_df))
```
