---
title: "Agencies Regex Table"
author: "Shirlyn Dong"
date: "`r Sys.Date()`"
format: 
  html:
    toc: true
vignette: >
  %\VignetteIndexEntry{Agencies}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = FALSE,
  warning = FALSE,
  message = FALSE,
  tidy = FALSE,
  fig.align = 'center',
  fig.path = "figures/agencies-",
  R.options = list(width = 120)
)

kable <- function(x){
  if (knitr::is_latex_output()) {
    head(x, 5) |>
      knitr::kable(booktabs = TRUE, format = 'latex') |>
      kableExtra::kable_styling(latex_options = c("striped", "scale_down", "HOLD_position"))
  } else {
    head(x, 5) |>
      knitr::kable(escape = TRUE) |>
      kableExtra::kable_styling()
  }
}
```

### Introduction

This vignette demonstrates how to extract sub agency names using a regex lookup table of agencies, sub agency names, sections, departments, subagency_nmae, and other_aliases The `regextable::extract()` workflow allows consistent identification of corporation names across datasets, even when multiple aliases or variants are used.

Install and load the package:
```{r, message=FALSE}
library(regextable)
library(kableExtra)
library(rvest)
library(googlesheets4)
library(purrr)
library(tibble)
library(stringr)
```

```{r}
load("/Users/shirl/Downloads/agencies_by_section.rda")
kable(head(agencies_regex))
```


```{r}
load("/Users/shirl/Downloads/body_parsed.rda")
kable(head(racialized_terms))
```


