---
title: "Corporations Regextable"
author: "Shirlyn Dong"
date: "`r Sys.Date()`"
format: 
  html:
    toc: true
vignette: >
  %\VignetteIndexEntry{Corporations}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = FALSE,
  warning = FALSE,
  message = FALSE,
  tidy = FALSE,
  fig.align = 'center',
  R.options = list(width = 120)
)

kable <- function(x, ...){
  if (knitr::is_latex_output()) {
    knitr::kable(x, booktabs = TRUE, format = 'latex', ...) |>
      kableExtra::kable_styling(latex_options = c("striped", "scale_down", "HOLD_position"))
  } else {
    knitr::kable(x, escape = TRUE, ...) |>
      kableExtra::kable_styling()
  }
}
```

### Introduction

This vignette demonstrates how to extract mentions of corporations using a regex lookup table of corporation names, aliases, tickers, and other metadata. The `regextable::extract()` workflow allows consistent identification of corporation names across datasets, even when multiple aliases or variants are used.

Install and load the package:
```{r, message=FALSE}
library(regextable)
library(kableExtra)
library(rvest)
library(pbapply)
library(purrr)
library(tibble)
library(stringr)
```

### Regex Table of Corporations

The `corporations_regex` table contains a sample of corporations, their common aliases, ticker symbols, and reference sources used for matching and standardization. The `aliases` column includes alternative names, abbreviations, or common variants for each corporation to ensure comprehensive matching.

```{r}
corporations_regex <- read.csv("/Users/shirl/Downloads/corporations_regex.csv", stringsAsFactors = FALSE)
kable(head(corporations_regex))
```

### Sample Dataset: Project 2025

The following dataset contains organizations and contributors involved in Project 2025. It is used to demonstrate matching and standardizing corporation names using the regex table.

```{r}
project_2025_url <- "https://raw.githubusercontent.com/judgelord/corporations/main/data/project_2025_coalition_and_contributors.rda"
tmp <- tempfile(fileext = ".rda")
download.file(project_2025_url, tmp, mode = "wb")
load(tmp)
unlink(tmp)
remove(tmp, project_2025_url)
kable(head(project_2025_coalition_and_contributors))
```

### Extracting Corporation Names

The `extract()` function searches the `organization` column of the dataset using the `patterns` column in the regex table. It returns the standardized corporation names while optionally removing acronym-only matches to reduce false positives.

```{r}
corp_df <- extract(data = project_2025_coalition_and_contributors,
                   col_name = "organization",
                   regex_table = corporations_regex,
                   data_return_cols = "organization",
                   remove_acronyms = TRUE,
                   use_ner = TRUE)
kable(head(corp_df))
```
