---
title: "Corporations Regextable"
author: "Shirlyn Dong"
date: "`r Sys.Date()`"
format: 
  html:
    toc: true
vignette: >
  %\VignetteIndexEntry{Corporations}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = FALSE,
  warning = FALSE,
  message = FALSE,
  tidy = FALSE,
  fig.align = 'center',
  fig.path = "figures/corporations",
  R.options = list(width = 120)
)

# Format kable tables
kable <- function(x){
  if (knitr::is_latex_output()) {
    head(x, 5) |>
      knitr::kable(booktabs = TRUE, format = 'latex') |>
      kableExtra::kable_styling(latex_options = c("striped", "scale_down", "HOLD_position"))
  } else {
    head(x, 5) |>
      knitr::kable(escape = TRUE) |>
      kableExtra::kable_styling()
  }
}
```

### Introduction

This vignette demonstrates how to extract mentions of corporations using a regex lookup table of corporation names, aliases, tickers, and other metadata. The `regextable::extract()` workflow allows consistent identification of corporation names across datasets, even when multiple aliases or variants are used.

Install and load the package:
```{r, message=FALSE}
library(regextable)
library(kableExtra)
library(rvest)
library(pbapply)
library(purrr)
library(tibble)
library(stringr)
```

### Regex Table of Corporations

The corporations_regex table contains corporations, their common aliases, ticker symbols, and reference sources used for matching and standardization. The aliases column includes alternative names, abbreviations, or common variants for each corporation to ensure comprehensive matching.

```{r}
corporations_regex <- read.csv("/Users/shirl/Downloads/corporations_crosswalk.csv", stringsAsFactors = FALSE)

# Clean aliases and create pattern

corporations_regex$aliases <- tolower(corporations_regex$aliases)
suffixes <- c(
  "\\binc\\b", "\\bcorp\\b", "\\bcorporation\\b",
  "\\bllc\\b", "\\blp\\b", "\\bltd\\b", "\\bincorporated\\b"
)
suffix_pattern <- paste0("(?:", paste(suffixes, collapse = "|"), ")")
corporations_regex$aliases <- gsub(
  paste0("[,.]?\\s*", suffix_pattern, "[.]?"),
  "",
  corporations_regex$aliases,
  perl = TRUE
)
corporations_regex$aliases <- gsub("[[:punct:]]+$", "", corporations_regex$aliases)
corporations_regex$aliases <- gsub("\\s+", " ", corporations_regex$aliases)
corporations_regex$aliases <- trimws(corporations_regex$aliases)
pbapply::pboptions(type = "none")
corporations_regex$pattern <- pbapply::pbsapply(corporations_regex$aliases, function(x){
  parts <- unlist(strsplit(x, "\\|"))
  parts <- trimws(parts)
  parts <- parts[nchar(parts) > 1]
  parts <- unique(parts)
  parts <- parts[order(nchar(parts), decreasing = TRUE)]
  paste(parts, collapse="|")
})
corporations_regex <- corporations_regex[nchar(corporations_regex$pattern) > 1, ]
corporations_regex$pattern <- paste0("\\b(?:", corporations_regex$pattern, ")\\b")

kable(head(corporations_regex))
```

### Data Corporations

The following dataset contains organizations and contributors involved in Project 2025. It is used to demonstrate matching and standardizing corporation names using the regex table.

```{r}
project_2025_url <- "https://raw.githubusercontent.com/judgelord/corporations/main/data/project_2025_coalition_and_contributors.rda"
tmp <- tempfile(fileext = ".rda")
download.file(project_2025_url, tmp, mode = "wb")
load(tmp)
kable(head(project_2025_coalition_and_contributors))
```

### Extracting Aliases from Corporation Crosswalk

The `extract()` function searches the `organization` column of the dataset using the `aliases` patterns in the regex table. It returns the standardized corporation names while optionally removing acronym-only matches to reduce false positives.

```{r}
corp_df <- extract(data = project_2025_coalition_and_contributors,
                   col_name = "organization",
                   regex_table = corporations_regex,
                   data_return_cols = "organization",
                   remove_acronyms = TRUE)

kable(head(corp_df))
```
