---
title: "Native American Tribes Regex Table"
author: "Shirlyn Dong"
date: "`r Sys.Date()`"
format: 
  html:
    toc: true
vignette: >
  %\VignetteIndexEntry{Native American Tribes}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = FALSE,
  warning = FALSE,
  message = FALSE,
  tidy = FALSE,
  fig.align = "center",
  R.options = list(width = 120)
)

kable <- function(x, ...) {
  if (knitr::is_latex_output()) {
    knitr::kable(x, booktabs = TRUE, format = "latex", ...) |>
      kableExtra::kable_styling(latex_options = c("striped", "scale_down", "HOLD_position"))
  } else {
    knitr::kable(x, escape = TRUE, ...) |>
      kableExtra::kable_styling()
  }
}
```

### Introduction

This vignette demonstrates how to use the `regextable` package to extract references to Native American tribes from text. It showcases the core `extract()` workflow using a lookup table of tribe names and name variants.

Install and load the package:
```{r, message=FALSE}
library(regextable)
library(kableExtra)
library(googledrive)
```

### Regex Table of Native American Tribes

The following table contains tribe names and regex patterns used for matching. Each row represents a tribe and includes possible spelling variations or alternate names used in text.

This lookup table includes the following columns:

- Name
- Strings
- Source
- Website
- Notes
- Type
- Emphasis

```{r}
googledrive::drive_deauth()

googledrive::drive_download(
  googledrive::as_id("1_946hi1zXeRvlGWIztrZDKJEcPM2jUY0"),
  path = "tribes_regex.csv",
  overwrite = TRUE
)

tribes_regex <- read.csv("tribes_regex.csv", stringsAsFactors = FALSE)
kable(head(tribes_regex))
```

### Gathering Sample Data

To demonstrate extraction, this vignette collects tribe names from the National Congress of American Indians (NCAI) Tribal Directory. The directory provides publicly available information about federally recognized tribes. The scraping example uses the rvest, dplyr, purrr, tibble, and stringr packages for HTML parsing and iteration. The code is included for reference only and is not evaluated in this vignette.

The following code can be run to scrape the live directory directly from the website:

```{r, eval = FALSE}
max_page <- read_html("https://www.ncai.org/tribal-directory") %>%
  html_elements(".Pagination_numberButton__vLhpm") %>%
  html_text() %>%
  as.numeric() %>%
  max(na.rm = TRUE)

all_pages <- paste0("https://www.ncai.org/tribal-directory/page/", 1:max_page)

tribes_df <- map_df(all_pages, function(url) {
  message("Scraping: ", url)
  html <- read_html(url)
  cards <- html %>% html_elements("article.TribeCard_tribeCard__UJcdx")

  map_df(cards, function(card) {
    tibble(
      Region = card %>% html_element(".TribeCard_regionLabel___OVFL")
        %>% html_text(trim = TRUE)
        %>% str_remove(" Region"),
      Tribe = card %>% html_element("h2") %>% html_text(trim = TRUE),
      Recognition = card %>% html_element(".TribeCard_federal__bQB0g")
        %>% html_text(trim = TRUE),
      District = card %>% html_element(".TribeCard_generic__MLwRU")
        %>% html_text(trim = TRUE)
        %>% str_remove("Congressional District ")
    )
  })
})
```

Live web scraping may take a long time to complete and is sensitive to changes in website structure. For reproducibility and convenience, a pre-scraped version of the dataset is available for download at: 

[Tribes data frame](https://drive.google.com/file/d/1vL_dI4SuvnxwIXr40X12uBoGSSDn_xL-/view)

For this example, a smaller sample of the full dataset is used:

```{r}
googledrive::drive_download(
  googledrive::as_id("1y4PwIYBdXoW6RmlGUeA9Qk3Josp0P0dl"),
  path = "tribes_df.csv",
  overwrite = TRUE
)

tribes_df <- read.csv("tribes_df.csv", stringsAsFactors = FALSE)
kable(head(tribes_df))
```

### Extracting Tribe Names from Directory

This example demonstrates how to use `regextable::extract()` to parse and standardize the Native American tribe names found within the scraped Tribal Directory data (`tribes_df`).

The `extract()` function searches the `Tribe` column using the regex patterns located in the `Strings` column of `tribes_regex`. When a match is found, the function returns the original tribe name along with metadata from the regex table, such as the source reference.

```{r}
tribe_directory_df <- extract(
  data = tribes_df,
  regex_table = tribes_regex,
  col_name = "Tribe",
  pattern_col = "Strings",
  data_return_cols = "Tribe",
  regex_return_cols = "Source"
)

kable(head(tribe_directory_df))
```

```{r cleanup, include = FALSE}
# Deletes the temporary CSVs so they don't clutter the project folder
unlink(c("tribes_regex.csv", "tribes_df.csv"))
```
