---
title: "Native American Tribes Regex Table"
author: "Shirlyn Dong"
date: "`r Sys.Date()`"
format: 
  html:
    toc: true
vignette: >
  %\VignetteIndexEntry{Native American Tribes}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = FALSE,
  warning = FALSE,
  message = FALSE,
  tidy = FALSE,
  fig.align = 'center',
  fig.path = "figures/native-tribes-",
  R.options = list(width = 120)
)

# Format kable tables
kable <- function(x){
  if (knitr::is_latex_output()) {
    head(x, 5) |>
      knitr::kable(booktabs = TRUE, format = 'latex') |>
      kableExtra::kable_styling(latex_options = c("striped", "scale_down", "HOLD_position"))
  } else {
    head(x, 5) |>
      knitr::kable(escape = TRUE) |>
      kableExtra::kable_styling()
  }
}
```

### Introduction

This vignette demonstrates how to extract mentions of Native American Tribes from public comment data using a regular expression (regex) lookup table of Native American Tribes and Groups. This approach is similar to the core `regextable::extract()` workflow but uses a specialized lookup table of tribe names and name variants. 

Install and load the package:
```{r, message=FALSE}
library(regextable)
library(kableExtra)
library(rvest)
```

### Native American Tribe Regex Table

This lookup table includes:
- Name
- Strings
- Source
- Website
- Notes
- Type
- Emphasis

```{r}
library(googlesheets4)
gs4_deauth()

tribes_regex <- read_sheet("1LaTgX3ffCrND5Nj1r7TzZqWvrit_clPGk2zIoNxo9Hg")
kable(head(tribes_regex))
```

### Data Agencies

```{r}
html <- read_html("https://www.ncai.org/tribal-directory")

tribe_names <- html %>% 
  html_nodes("h2.TribeCard_regionTitle__8Sffj") %>% 
  html_text(trim = TRUE)

regions <- html %>%
  html_nodes(".TribeCard_regionLabel___OVFL") %>%
  html_text(trim = TRUE)

congressional_district <- html %>%
  html_nodes("li.TribeCard_tagItem__UjDJv.TribeCard_generic_MLwRU") %>%
  html_text(trim = TRUE)

recognition <- html %>%
  html_nodes("li.TribeCard_tagItem__UjDJv.TribeCard_federal__bQB0g") %>%
  html_text(trim = TRUE)

leader <- html %>%
  html_node(".TribeCard_columns__jeDha") %>%
  html_node("p") %>%
  html_text(trim = TRUE)

tribes_df <- data.frame(
  Tribe = tribe_names,
  Regions = regions,
  Congressional_District = congressional_district,
  stringsAsFactors = FALSE
)

kable(head(tribes))
```
