---
title: "Native American Tribes Regex Table"
author: "Shirlyn Dong"
date: "`r Sys.Date()`"
format: 
  html:
    toc: true
vignette: >
  %\VignetteIndexEntry{Native American Tribes}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = FALSE,
  warning = FALSE,
  message = FALSE,
  tidy = FALSE,
  fig.align = 'center',
  fig.path = "figures/native-tribes-",
  R.options = list(width = 120)
)

# Format kable tables
kable <- function(x){
  if (knitr::is_latex_output()) {
    head(x, 5) |>
      knitr::kable(booktabs = TRUE, format = 'latex') |>
      kableExtra::kable_styling(latex_options = c("striped", "scale_down", "HOLD_position"))
  } else {
    head(x, 5) |>
      knitr::kable(escape = TRUE) |>
      kableExtra::kable_styling()
  }
}
```

### Introduction

This vignette demonstrates how to extract mentions of Native American Tribes from public comment data using a regular expression (regex) lookup table of Native American Tribes and Groups. This approach is similar to the core `regextable::extract()` workflow but uses a specialized lookup table of tribe names and name variants. 

Install and load the package:
```{r, message=FALSE}
library(regextable)
library(kableExtra)
library(httr2)
library(base64enc)
```

### Native American Tribe Regex Table

This lookup table includes:
- Name
- Strings
- Source
- Website
- Notes
- Type
- Emphasis

```{r}
library(googlesheets4)
gs4_deauth()

tribes_regex <- read_sheet("1LaTgX3ffCrND5Nj1r7TzZqWvrit_clPGk2zIoNxo9Hg")
kable(head(tribes_regex))
```

### Data Agencies
```{r load_github_rda, include=TRUE}
get_github_rda <- function(owner, repo, path) {
  url <- sprintf(
    "https://api.github.com/repos/%s/%s/contents/%s",
    owner, repo, path
  )
  
  resp <- httr2::request(url) |> httr2::req_perform()
  json <- httr2::resp_body_json(resp)
  
  raw_file <- base64enc::base64decode(json$content)
  
  tmp <- tempfile(fileext = ".rda")
  writeBin(raw_file, tmp)
  load(tmp, envir = .GlobalEnv)
}

get_github_rda(
  owner = "judgelord",
  repo = "2025",
  path = "data/body_text_clean.rda"
)
kable(head(body_text_clean))
```
