---
title: "Native American Tribes Regex Table"
author: "Shirlyn Dong"
date: "`r Sys.Date()`"
format: 
  html:
    toc: true
vignette: >
  %\VignetteIndexEntry{Native American Tribes}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = FALSE,
  warning = FALSE,
  message = FALSE,
  tidy = FALSE,
  fig.align = 'center',
  R.options = list(width = 120)
)

kable <- function(x, ...){
  if (knitr::is_latex_output()) {
    knitr::kable(x, booktabs = TRUE, format = 'latex', ...) |>
      kableExtra::kable_styling(latex_options = c("striped", "scale_down", "HOLD_position"))
  } else {
    knitr::kable(x, escape = TRUE, ...) |>
      kableExtra::kable_styling()
  }
}
```

### Introduction

This vignette demonstrates how to extract references to Native American tribes using a regex lookup table. This approach is similar to the core `regextable::extract()` workflow but uses a specialized lookup table of tribe names and name variants. 

Install and load the package:
```{r, message=FALSE}
library(regextable)
library(kableExtra)
library(rvest)
library(purrr)
library(tibble)
library(stringr)
```

### Regex Table of Native American Tribes

The following table contains tribe names and regex patterns used for matching. Each row represents a tribe and includes possible spelling variations or alternate names used in text.

This lookup table includes:
- Name
- Strings
- Source
- Website
- Notes
- Type
- Emphasis

```{r}
tribes_regex <- read.csv("/Users/shirl/Downloads/Native_American_Tribes_Regex_Table.csv", stringsAsFactors = FALSE)
kable(head(tribes_regex))
```

### Extracting Tribe Names from Directory

To demonstrate extraction, this vignette collects tribe names from the National Congress of American Indians (NCAI) Tribal Directory. The directory provides publicly available information about federally recognized tribes.

```{r}
max_page <- read_html("https://www.ncai.org/tribal-directory") %>%
  html_elements(".Pagination_numberButton__vLhpm") %>%
  html_text() %>%
  as.numeric() %>%
  max(na.rm = TRUE)

all_pages <- paste0("https://www.ncai.org/tribal-directory/page/", 1:max_page)

tribes_df <- map_df(all_pages, function(url) {
  
  message("Scraping: ", url)
  html <- read_html(url)
  cards <- html %>% html_elements("article.TribeCard_tribeCard__UJcdx")
  
  map_df(cards, function(card) {
    tibble(
      Region = card %>% html_element(".TribeCard_regionLabel___OVFL") 
                    %>% html_text(trim = TRUE) 
                    %>% str_remove(" Region"),
      Tribe = card %>% html_element("h2") %>% html_text(trim = TRUE),
      Leader = card %>% html_element("section:nth-of-type(1) p:nth-of-type(1)") 
                    %>% html_text(trim = TRUE),
      Tel = card %>% html_element(xpath = ".//p[strong[contains(., 'Tel:')]]") 
                 %>% html_text(trim = TRUE) 
                 %>% str_remove("Tel: "),
      Fax = card %>% html_element(xpath = ".//p[strong[contains(., 'Fax:')]]") 
                 %>% html_text(trim = TRUE) 
                 %>% str_remove("Fax: "),
      Address = card %>% html_element("section:nth-of-type(2)") %>% html_text2(),
      Recognition = card %>% html_element(".TribeCard_federal__bQB0g") 
                         %>% html_text(trim = TRUE),
      District = card %>% html_element(".TribeCard_generic__MLwRU") 
                      %>% html_text(trim = TRUE)
                      %>% str_remove("Congressional District ")
    )
  })
})

kable(head(tribes_df))
```

### Extracting Tribe Names from Directory

This example demonstrates how to use `regextable::extract()` to extract Native American Tribe names in the Tribal Directory data (`tribes_df`). The `extract()` function searches the `Tribe` column using the regex patterns in the `Strings` column of `tribes_regex`. When a match is found, the function returns the original tribe name along with metadata from the regex table, such as the source reference.

```{r}
tribe_directory_df <- extract(
    data = tribes_df,
    regex_table = tribes_regex,
    col_name = "Tribe",
    pattern_col = "Strings",
    data_return_cols = "Tribe",
    regex_return_cols = "Source"
)

kable(head(tribe_directory_df))
```
