---
title: "Native American Tribes Regex Table"
author: "Shirlyn Dong"
date: "`r Sys.Date()`"
format: 
  html:
    toc: true
vignette: >
  %\VignetteIndexEntry{Native American Tribes}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = FALSE,
  warning = FALSE,
  message = FALSE,
  tidy = FALSE,
  fig.align = 'center',
  fig.path = "figures/native-tribes-",
  R.options = list(width = 120)
)

# Format kable tables
kable <- function(x){
  if (knitr::is_latex_output()) {
    head(x, 5) |>
      knitr::kable(booktabs = TRUE, format = 'latex') |>
      kableExtra::kable_styling(latex_options = c("striped", "scale_down", "HOLD_position"))
  } else {
    head(x, 5) |>
      knitr::kable(escape = TRUE) |>
      kableExtra::kable_styling()
  }
}
```

### Introduction

This vignette demonstrates how to extract mentions of Native American Tribes from public comment data using a regular expression (regex) lookup table of Native American Tribes and Groups. This approach is similar to the core `regextable::extract()` workflow but uses a specialized lookup table of tribe names and name variants. 

Install and load the package:
```{r, message=FALSE}
library(regextable)
library(kableExtra)
library(rvest)
```

### Native American Tribe Regex Table

This lookup table includes:
- Name
- Strings
- Source
- Website
- Notes
- Type
- Emphasis

```{r}
library(googlesheets4)
gs4_deauth()

tribes_regex <- read_sheet("1LaTgX3ffCrND5Nj1r7TzZqWvrit_clPGk2zIoNxo9Hg")
kable(head(tribes_regex))
```

### Data Agencies

```{r}

cards <- read_html("https://www.ncai.org/tribal-directory") %>% 
  html_elements("article.TribeCard_tribeCard__UJcdx")

tribes_df <- map_df(cards, function(card) {
  tibble(
    Region = card %>% html_element(".TribeCard_regionLabel___OVFL") %>% html_text(trim = TRUE),
    Tribe = card %>% html_element("h2") %>% html_text(trim = TRUE),
    Leader = card %>% html_element("section:nth-of-type(1) p:nth-of-type(1)") 
                  %>% html_text(trim = TRUE),
    
    Tel = card %>% html_element(xpath = ".//p[strong[contains(., 'Tel:')]]") 
               %>% html_text(trim = TRUE) %>% str_remove("Tel: "),
    Fax = card %>% html_element(xpath = ".//p[strong[contains(., 'Fax:')]]") 
               %>% html_text(trim = TRUE) %>% str_remove("Fax: "),
    
    Address  = card %>% html_element("section:nth-of-type(2)") %>% html_text2(),
    
    Recognition = card %>% html_element(".TribeCard_federal__bQB0g") %>% html_text(trim = TRUE),
    District = card %>% html_element(".TribeCard_generic_MLwRU") %>% html_text(trim = TRUE)
  )
})

kable(head(tribes_df))
```
