---
title: "regextable"
author: "Shirlyn Dong"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{regextable}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = FALSE,
  warning = FALSE,
  message = FALSE,
  tidy = FALSE,
  fig.align = 'center',
  fig.path = "man/figures/vignette-",
  R.options = list(width = 120)
)

str_fix <- function(x){
  stringr::str_replace_all(x, "\\$", "\\\\$") |> 
    stringr::str_remove_all("`|\\\\bf|\\\\br")
}

# Format kable tables
kable <- function(...){
  if (knitr::is_latex_output()){
    head(..., 25) |>
      knitr::kable(booktabs = TRUE, format = 'latex') |>
      kableExtra::kable_styling(latex_options = c("striped", "scale_down", "HOLD_position"))
  } else {
    head(..., 50) |>
      dplyr::mutate(dplyr::across(everything(), str_fix)) |>
      knitr::kable(escape = TRUE) |>
      kableExtra::kable_styling()
  }
}
```


### Introduction

`regextable` provides a flexible and efficient framework for extracting structured information from text using pattern-based regular expressions. The package is optimized to handle large datasets with thousands of patterns using a "shrinking pool" algorithm, which stops searching rows once a match is found, improving speed.

Install and load the package:
```{r, eval= FALSE}
devtools::install_github("judgelord/regextable")
```
```{r, message=FALSE}
library(regextable)
```

## Data

For demonstration, we use two included datasets:

- `members`: A lookup table of regex patterns for member names.
- `cr2007_03_01`: A sample text dataset to search.

```{r}
data("members")
kable(head(members))

data("cr2007_03_01")
kable(head(cr2007_03_01))
```

### Text Cleaning

`extract()` automatically cleans text by default, so the user does not need to call manually. Cleaning standardizes spacing, punctuation, and capitalization, which helps regex patterns match more reliably.

Example of `clean_text()`:
```{r}
text <- "  HELLO---WORLD  "
clean_text(text)
```

## Basic Extraction

The simplest use of `extract()`:

```{r}
result <- extract(
  data = cr2007_03_01,
  col_name = "speaker",
  regex_table = members,
  return_cols = c("speaker", "pattern"),
  regex_return_cols = c("icpsr")
)

kable(head(result))
```

Explanation:
- `data`: the text dataset to search.
- `col_name`: which column contains the text.
- `regex_table`: the lookup table of patterns.
- `return_cols`: columns from data to include in the result.
- `regex_return_cols`: additional columns from the pattern table to attach.
Each row in the output corresponds to a detected match, and includes both the original text and the matching pattern.
---

## Advanced Usage

`extract()` can also filter data by date, remove all-uppercase patterns (like acronyms), and select specific output columns. This is useful for more controlled extraction.

```{r, eval = FALSE, include= FALSE}
result_advanced <- extract(
  data = cr2007_03_01,
  col_name = "speaker",
  regex_table = members,
  date_col = "date",
  date_start = "2007-01-01",
  date_end = "2007-12-31",
  remove_acronyms = TRUE,
  return_cols = c("speaker", "pattern"),
  regex_return_cols = c("icpsr")
)

kable(head(result_advanced))
```

Explanation:
- `date_col`, date_start, date_end: filter rows by date.
- `remove_acronyms`: skip patterns like "NASA" or "USA".
You can combine these filters with any subset of columns for flexible outputs.
---

### How the Matching Works

Internally, `extract()` uses a shrinking pool algorithm:

1. Start with all text rows.
2. Check each regex pattern against unmatched rows.
3. When a match is found, remove that row from further checks.
4. Continue until all patterns have been applied.

This approach greatly improves performance on large datasets because rows stop being checked once a match is found.

### Summary
- `regextable` is a tool for extracting structured data from unstructured text.
- Use the included datasets to get started or supply your own lookup tables.
- `extract()` automatically handles text cleaning and efficient matching.
- Optional parameters allow advanced control over filtering and output.
