---
title: "regextable"
author: "Shirlyn Dong"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{regextable}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = FALSE,
  warning = FALSE,
  message = FALSE,
  tidy = FALSE,
  fig.align = 'center',
  fig.path = "man/figures/vignette-",
  R.options = list(width = 120)
)

# Format kable tables
kable <- function(x){
  if (knitr::is_latex_output()) {
    head(x, 25) |>
      knitr::kable(booktabs = TRUE, format = 'latex') |>
      kableExtra::kable_styling(latex_options = c("striped", "scale_down", "HOLD_position"))
  } else {
    head(x, 50) |>
      knitr::kable(escape = TRUE) |>
      kableExtra::kable_styling()
  }
}
```


### Introduction

`regextable` extracts regex-based pattern matches from a data frame or character vector using a pattern lookup table, returning the matched pattern, the exact text match, an internal row identifier, and all original columns from the input data unless specified, with optional additional columns from the pattern table. Performance on large datasets is improved by stopping the search for a row once a match is found.

Install and load the package:
```{r, eval= FALSE}
devtools::install_github("judgelord/regextable")
```
```{r, message=FALSE}
library(regextable)
```

## Data

For demonstration, we use two included datasets:

- `members`: A lookup table of regex patterns for member names.
- `cr2007_03_01`: A sample text dataset to search.

```{r}
data("members")
kable(head(members))

data("cr2007_03_01")
kable(head(cr2007_03_01))
```

### Text Cleaning

`extract()` cleans text by default, so the user does not need to call it manually. Cleaning standardizes spacing, punctuation, and capitalization, which helps regex pattern matching.

Example of `clean_text()`:
```{r}
text <- "  HELLO---WORLD  "
clean_text(text)
```

## Basic Extraction

The simplest use of `extract()`:

```{r}
result <- extract(
  data = cr2007_03_01,
  regex_table = members,
  return_cols = c("text"),
  regex_return_cols = c("icpsr")
)

kable(head(result))
```

Explanation:
- `data`: the text dataset to search.
- `col_name`: which column contains the text.
- `regex_table`: the lookup table of patterns.
- `return_cols`: columns from data to include in the result.
- `regex_return_cols`: additional columns from the pattern table to attach.
Each row in the output corresponds to a detected match, and includes both the original text and the matching pattern.
---

## Advanced Usage

`extract()` can also filter data by date, remove acronyms (all-uppercase patterns with 2+ characters), and select specific output columns. This is useful for more controlled extraction.

```{r, eval = FALSE, include= FALSE}
result_advanced <- extract(
  data = cr2007_03_01,
  regex_table = members,
  date_col = "date",
  date_start = "2007-01-01",
  date_end = "2007-12-31",
  remove_acronyms = TRUE,
  return_cols = c("text"),
  regex_return_cols = c("icpsr")
)

kable(head(result_advanced))
```

Explanation:
- `date_col`, date_start, date_end: filter rows by date.
- `remove_acronyms`: skip patterns like "NASA" or "USA".
You can combine these filters with any subset of columns for flexible outputs.
---

### How the Matching Works

Internally, `extract()` uses a shrinking pool algorithm:

1. Start with all text rows.
2. Check each regex pattern against unmatched rows.
3. When a match is found, remove that row from further checks.
4. Continue until all patterns have been applied.

This approach improves performance on large datasets because rows stop being checked once a match is found.

### Summary
- `regextable` is a tool for extracting data from text.
- Use the included datasets to get started or supply your own lookup tables.
- `extract()` by default handles text cleaning and efficient matching.
- Optional parameters allow advanced control over filtering and output.
