---
title: "regextable"
author: "Shirlyn Dong"
date: "`r Sys.Date()`"
format: html
vignette: >
  %\VignetteIndexEntry{regextable}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = FALSE,
  warning = FALSE,
  message = FALSE,
  tidy = FALSE,
  fig.align = 'center',
  R.options = list(width = 120)
)

kable <- function(x, ...){
  if (knitr::is_latex_output()) {
    knitr::kable(x, booktabs = TRUE, format = 'latex', ...) |>
      kableExtra::kable_styling(latex_options = c("striped", "scale_down", "HOLD_position"))
  } else {
    knitr::kable(x, escape = TRUE, ...) |>
      kableExtra::kable_styling()
  }
}
```

### Introduction

`regextable` extracts regex-based pattern matches from a data frame or character vector using a pattern lookup table. For each input row, all matching patterns are returned, along with the matched substring, an internal row identifier, and additional columns specified in `data_return_cols` and `regex_return_cols`. Optional metadata from the pattern table can also be included. Multiple rows may be returned for a single text if it matches multiple patterns.

The function also supports additional parameters such as `do_clean_text`, `unique_match`, `verbose`, `cl`, and optional Named Entity Recognition validation via `use_ner.` These can be used to control text cleaning, limit matches per row, show progress messages, or speed up processing on large datasets.

Install and load the package:
```{r, message=FALSE}
library(regextable)
library(kableExtra)
```

### Data

For demonstration, we use two included datasets:

- `members`: A lookup table of regex patterns for member names.
- `cr2007_03_01`: A sample text dataset to search.

```{r}
data("members")
kable(members)

data("cr2007_03_01")
kable(cr2007_03_01)
```

### Text Cleaning

`extract()` performs basic text cleaning by default - standardizing spaces, punctuation, and capitalization—to improve pattern matching. This can be disabled by setting `do_clean_text = FALSE`. Cleaning standardizes spacing, punctuation, and capitalization, which helps regex pattern matching.

Example of `clean_text()`:
```{r}
text <- "  HELLO---WORLD  "
clean_text(text)
```

### Basic Extraction

The simplest use of `extract()`:

```{r}
result <- regextable::extract(
  data = cr2007_03_01,
  regex_table = members,
  data_return_cols = c("text"),
  regex_return_cols = c("icpsr")
)

kable(head(result))
```

Explanation:
- `data`: the text dataset to search.
- `col_name`: which column contains the text.
- `regex_table`: the lookup table of patterns.
- `data_return_cols`: additional columns from `data` to include in the result.
- `regex_return_cols`: additional columns from the pattern table to attach.
Each row in the output corresponds to a detected match, and includes both the original text and the matching pattern.
---

### Advanced Usage

`extract()` can also filter data by date, remove acronyms (all-uppercase patterns with 2+ characters), and select specific output columns. This is useful for more controlled extraction.

```{r, eval = FALSE}
result_advanced <- extract(
  data = cr2007_03_01,
  regex_table = members,
  date_col = "date",
  date_start = "2007-01-01",
  date_end = "2007-12-31",
  remove_acronyms = TRUE,
  data_return_cols = c("text"),
  regex_return_cols = c("icpsr")
)

kable(head(result_advanced))
```

Explanation:
- `date_col`, `date_start`, `date_end`: filter rows by date.
- `remove_acronyms`: skip patterns like "NASA" or "USA".
You can combine these filters with any subset of columns for flexible outputs.
---

### Named Entity Recognition (NER) Validation

`extract()` optionally supports validation using Named Entity Recognition (NER).
When `use_ner = TRUE`, the function uses the `spacyr` package to confirm that
regex matches correspond to actual named entities in the text.

Two arguments used for this behavior:

- `use_ner`: Logical; if `TRUE`, validates matches using `spacyr` Named Entity Recognition.
- `ner_entity_types`: Character vector specifying which entity types to retain.
  For example, `"ORG"` keeps only organization entities.

Example:

```{r, eval = FALSE}
spacyr::spacy_initialize()

result_ner <- regextable::extract(
  data = cr2007_03_01,
  regex_table = members,
  use_ner = TRUE,
  ner_entity_types = c("PERSON"),
  data_return_cols = "text",
  regex_return_cols = "icpsr"
)
```

### Parallel Matching

`extract()` supports parallel processing via the `cl` parameter:

```{r, eval=FALSE}
library(parallel)
clust <- makeCluster(2)
result_parallel <- regextable::extract(
  data = cr2007_03_01,
  regex_table = members,
  cl = clust,
  data_return_cols = c("text"),
  regex_return_cols = c("icpsr")
)
stopCluster(clust)
kable(head(result_parallel))
```

### Summary
- `regextable` is a tool for extracting data from text.
- Use the included datasets to get started or supply your own lookup tables.
- `extract()` by default handles text cleaning and efficient matching.
- Optional parameters allow advanced control over filtering and output.
