---
title: "regextable"
author: "Shirlyn Dong"
date: "`r Sys.Date()`"
format: html
vignette: >
  %\VignetteIndexEntry{regextable}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = FALSE,
  warning = FALSE,
  message = FALSE,
  tidy = FALSE,
  fig.align = 'center',
  R.options = list(width = 120)
)

kable <- function(x, ...){
  if (knitr::is_latex_output()) {
    knitr::kable(x, booktabs = TRUE, format = 'latex', ...) |>
      kableExtra::kable_styling(latex_options = c("striped", "scale_down", "HOLD_position"))
  } else {
    knitr::kable(x, escape = TRUE, ...) |>
      kableExtra::kable_styling()
  }
}
```

### Introduction

The `regextable` package extracts regex-based pattern matches from a data frame or character vector using a pattern lookup table. For each input row, all matching patterns are returned, along with the matched substring, an internal row identifier, and additional columns specified in `data_return_cols` and `regex_return_cols`. Optional metadata from the pattern table can also be included. Multiple rows may be returned for a single text if it matches multiple patterns.

The function also supports additional parameters such as `do_clean_text`, `unique_match`, `verbose`, `cl`, and optional Named Entity Recognition validation via `use_ner`. These can be used to control text cleaning, optimize performance, show progress messages, or speed up processing on large datasets.

Install and load the package:
```{r, message=FALSE}
library(regextable)
library(kableExtra)
```

### Example Data

For demonstration, two included datasets are used:
- `members`: A lookup table of regex patterns for members of Congress.
- `cr2007_03_01`: A sample text dataset from the Congressional Record.

```{r}
data("members")
kable(head(members))

data("cr2007_03_01")
kable(head(cr2007_03_01))
```

### Text Cleaning

By default, `extract()` performs basic text cleaning to improve pattern matching. It converts text to lowercase, removes specific punctuation (`+`, `—`, `!`, `?`, `:`, `;`), replaces line breaks, tabs, periods, and dashes with spaces, and normalizes commas and excess whitespace. This can be disabled by setting `do_clean_text = FALSE`.

Example of `clean_text()`:
```{r}
text <- "  HELLO---WORLD  "
clean_text(text)
```

### Basic Extraction

The simplest use of extract():
```{r}
result <- regextable::extract(
  data = cr2007_03_01,
  regex_table = members,
  data_return_cols = c("text"),
  regex_return_cols = c("icpsr")
)

kable(head(result))
```

Function Arguments:
- `data`: The text dataset to search.
- `col_name`: Which column contains the text (defaults to `"text"`). Note: If a simple character vector is passed instead of a data frame, it is internally converted, and this argument is ignored.
- `regex_table`: The lookup table of patterns.
- `data_return_cols`: Additional columns from `data` to include in the result.
- `regex_return_cols`: Additional columns from the pattern table to attach.

Each row in the output corresponds to a detected match, and includes both the original text and the matching pattern.
---

### Advanced Usage

By default, `extract()` evaluates every pattern against every text entry. If it is known that each row contains at most one match, setting `unique_match = TRUE` utilizes a faster algorithm that stops searching a row once the first match is found. This improves execution speed on large datasets.

```{r, eval = FALSE}
result_advanced <- extract(
  data = cr2007_03_01,
  regex_table = members,
  unique_match = TRUE,
  date_col = "date",
  date_start = "2007-01-01",
  date_end = "2007-12-31",
  remove_acronyms = TRUE,
  data_return_cols = c("text"),
  regex_return_cols = c("icpsr")
)

kable(head(result_advanced))
```

Function Arguments:
- `unique_match`: If TRUE, utilizes a faster algorithm that stops searching a row after the first match.
- `date_col`, `date_start`, `date_end`: Filter rows by date before matching.
- `remove_acronyms`: Skip patterns consisting entirely of uppercase letters (e.g., "NASA" or "USA").

These filters can be combined with any subset of columns for flexible outputs.
---

### Named Entity Recognition (NER) Validation

`extract()` optionally supports validation using Named Entity Recognition (NER). When `use_ner = TRUE`, the function uses the `spacyr` package to confirm that regex matches correspond to actual named entities in the text.

Function Arguments:
- `use_ner`: Logical; if `TRUE`, validates matches using `spacyr` Named Entity Recognition.
- `ner_entity_types`: Character vector specifying which entity types to retain.
  For example, `"ORG"` keeps only organization entities.

Example:
```{r, eval = FALSE}
spacyr::spacy_initialize()

result_ner <- regextable::extract(
  data = cr2007_03_01,
  regex_table = members,
  use_ner = TRUE,
  ner_entity_types = c("PERSON"),
  data_return_cols = "text",
  regex_return_cols = "icpsr"
)
```

### Parallel Matching

`extract()` supports parallel processing via the `cl` parameter, which is recommended for large datasets or extensive lookup tables.

```{r, eval=FALSE}
library(parallel)
clust <- makeCluster(2)
result_parallel <- regextable::extract(
  data = cr2007_03_01,
  regex_table = members,
  cl = clust,
  data_return_cols = c("text"),
  regex_return_cols = c("icpsr")
)
stopCluster(clust)
kable(head(result_parallel))
```

### Summary
- `regextable` is a tool for extracting data from text using lookup tables.
- `extract()` handles text cleaning and efficient matching by default.
- Advanced performance parameters, such as `unique_match` and `cl`, allow for significant speed improvements on large text.
- Optional parameters allow advanced control over filtering, NER validation, and output structure.